---
title: "Fledglings"
author: "Mara Bodenmann"
date: today
toc: true
abstract-title: "Research question:"
abstract: "How do environmental stressors and resource availability influence reproductive success?"
format: pdf
#  html:
#    embed-resources: true #single HTML file with no supportive directory
#    css: ../../styles.css
#    number-sections: true
#    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false #this prevent warning messages being rendered
  message: false
---

## Prepare Session

Import libraries

```{r}
#| label: load-libraries
#| code-fold: true

  library(pastecs)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)
  library(tidyverse)

  library(knitr) 
```

Import the data

```{r}
#| label: import-data
#| 

clean <- read_csv("data/processed/clean.CSV")

```

Modify the data

```{r}
#| label: modify-data

#deal with NA--> remove the row with this NA
fledglings<- clean%>%
  select(n_fledglings, insect_abundance, env_noise)%>%
  na.omit() #removes rows with na, can specify column: filter(!is.na(column_name))

```

Here the flags and the row with the NA was removed.

## Explore the data

### Numerically

```{r}
#| label: tbl-numerical-exploration-of-data
#| tbl-cap: "Numerical exploration of the data"

numexp<-stat.desc(fledglings,norm = T)

kable(numexp, digits=1)
```

### Graphically

```{r}
#| label: fig-graphical-exploration
#| fig-cap: "Raw data to visualize the relationship between successfully raised fledglings per nest, A: Insect abundance close to the nest and B: the environmental ambient noise."
#| fig-width: !expr 15/2.54
#| fig-height: !expr 9/2.54
#| code-fold: true
expgraph<-ggarrange(
  ggplot(fledglings, aes(x=insect_abundance,y=n_fledglings))+
    geom_point()+
    geom_smooth(method= glm, method.args= list(family= "poisson"),
                  col= "black")+
    labs(x=expression("Insect abundance in (g/m" ^2*")"),
         y="Number of succesfully raised fledglings per nest",
         title="A"),
  ggplot(fledglings, aes(x=env_noise,y=n_fledglings))+
    geom_point()+
    geom_smooth(method= glm, method.args= list(family= "poisson"),
                  col= "black")+
    labs(x="Environmental ambient noise (db)",
         y="Number of succesfully raised fledglings per nest",
         title="B")
)
expgraph
```

## Fit the models

Fitting different models to see which one is most suitable for the data

```{r}
#| label: fitting-the-models
mod1<-glm(n_fledglings~ 1, family= "poisson", data=fledglings)
mod2<-glm(n_fledglings~ insect_abundance + env_noise , family= "poisson",
          data=fledglings)
mod3<-glm(n_fledglings~ insect_abundance * env_noise , family= "poisson",
          data=fledglings)
```

### Evaluate the models 

Choose which one fits the best.

```{r}
#| label: tbl-evaluate-the-models
#| tbl-cap: "A table showing he evaluation of the models 1-3"
#| code-fold: true

anv.models<-anova(mod1,mod2,mod3)
kable(anv.models)
```

The model I choose is model 2 where there is no interaction between the insect abundance and the environment noise. Because with the interaction, the model was not significantly better than the one without the interaction.

## Validate the model

Validate whether I should proceed with the model

First I have to check for overdispersion because it is a glm!

```{r}
#| label: overdispersion
#| code-fold: true

check_overdispersion(mod2)

```

There was no overdispersion detected, we can continue with checking the model.

```{r}
#| label: fig-diagnostics-on-model
#| code-fold: true
#| fig-cap: "Model diagnostic "
#| fig-width: !expr 25/2.54
#| fig-height: !expr 29/2.54
check_model(mod2)

```

This also looks alright, I can continue with interpreting the model.

## Interpreting the model

```{r}
#| label: All-the-things-I-want-to-reference
#| code-fold: true

#coefficients
coef.table<-round(summary(mod2)$coefficients,3)

#chi^2 of my model
null.chi<-anova(mod1,mod2)

#get r squared adjusted
r.2<- round(attributes(r.squaredLR(mod2))$adj.r.squared, 3)

#prepare the predictions plot
pred.mod<-predict_response(mod2)

#make the plot
pred.plot <-ggarrange(
  ggplot(pred.mod$insect_abundance, aes(x,predicted))+
    geom_point(data=fledglings,aes(x=insect_abundance,y=n_fledglings))+
    geom_ribbon(aes(ymin= conf.low, ymax= conf.high), alpha=0.2)+
    geom_line()+
    labs(x=expression("Insect abundance in (g/m" ^2*")"),
         y="Number of succesfully raised fledglings per nest", 
         title = "A"),
  ggplot(pred.mod$env_noise, aes(x,predicted))+
    geom_point(data=fledglings,aes(x=env_noise,y=n_fledglings))+
    geom_ribbon(aes(ymin= conf.low, ymax= conf.high), alpha=0.2)+
    geom_line()+
    labs(x="Envronmental ambient noise (db)",
         y="Number of succesfully raised fledglings per nest",
         title = "B"),
  ncol=2,nrow=1
  
)
```

```{r}
#| label: tbl-glm
#| tbl-cap: !expr paste("Standard errors and statistical significance of the parameters estimated by a poisson GLM.", "($R^2$=", r.2, ", $\\chi^2_{(", null.chi[2, 3], ",", null.chi[2, 1], ")}$=", round(null.chi[2,4], 3), ", $p$", format.pval(null.chi[2, 5], eps = 0.001), ")")
#| echo: false

kable(coef.table)
```

A poisson Generalized Linear Model (*R*^2^= `r r.2` , `r paste("$\\chi^2_{(", null.chi[2,3], ",", null.chi[2,1], ")}$")`= `r round(null.chi[2, 4], 3)` , *p*`r format.pval(null.chi[2, 5], eps= 0.001)`; @tbl-glm) performed significantly better than a null-model.

The model generated moderate compelling support for the hypothesized effect of insect abundance on successfully raised fledglings per nest ($\beta_{Insect Abundance}$= `r round(coef.table[2, 1], 3)` , *SE*= `r round(coef.table[2, 2], 3)` , *z*= `r round(coef.table[2, 3], 3)`, *p*= `r round(coef.table[2,  4], 3)`) .

However no compelling support was generated by the model for the effect of environmental ambient noise on successfully raised fledglings per nest ($\beta_{Environmental AmbientNoise}$= `r round(coef.table[3, 1], 3)` , *SE*= `r round(coef.table[3, 2], 3)` , *z*= `r round(coef.table[3, 3], 3)`, *p*= `r round(coef.table[3,  4], 3)`) Thess findings is corroborated by @fig-predict , in which we see that, as the insect abundance increases the number of fledglings increases.

```{r}
#| label: fig-predict
#| fig-cap: "Raw data and model predictions to visualize the relationship between successfully raised fledglings per nest, A: Insect abundance close to the nest and B: the environmental ambient noise."
#| fig-width: !expr 15/2.54
#| fig-height: !expr 9/2.54
#| echo: false

# Plot predictions
  pred.plot
```

## Store session information

```{r}
#| label: store-session-info
#| code-fold: true

  version<- sessionInfo()


```
